---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---

```{r, echo = TRUE}
knitr::opts_chunk$set(warning=FALSE)
```
## Loading and preprocessing the data
```{r, echo = TRUE}
data <- read.csv("activity.csv")
dim(data)
head(data)
str(data)
```


## What is mean total number of steps taken per day?
```{r, echo = TRUE}
total_steps <- with(data, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(total_steps) <- c("date","steps")
hist(total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "red")
```


Here is the mean of the total number of steps taken per day
```{r, echo = TRUE}
mean(total_steps$steps)
```


Here is the median of the total number of steps taken per day
```{r, echo = TRUE}
median(total_steps$steps)
```


## What is the average daily activity pattern?
```{r, echo = TRUE}
avg_daily <- with(data, aggregate(data$steps, by = list(data$interval), FUN = mean, na.rm = TRUE))
names(avg_daily) <- c("interval", "mean")
plot(avg_daily$interval, avg_daily$mean, type = "l", col = "red", main = "Average number of steps per interval", xlab = "Intervals", ylab = "Average number of steps", lwd = 2)
```


## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo = TRUE}
sum(is.na(data$steps))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval.
```{r, echo = TRUE}
imputed_steps <- avg_daily$mean[match(data$interval, avg_daily$interval)]
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo = TRUE}
activity_imputed <- transform(data, steps = ifelse(is.na(data$steps), yes = imputed_steps, no = data$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r, echo = TRUE}
hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```

Here is the mean of the total number of steps taken per day:
```{r, echo = TRUE}
mean(total_steps_imputed$daily_steps)
```

Here is the median of the total number of steps taken per day:
```{r, echo = TRUE}
median(total_steps_imputed$daily_steps)
```

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r, echo = TRUE}
data$date <- as.Date(strptime(data$date, format="%Y-%m-%d"))
data$datetype <- sapply(data$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```

Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r, echo = TRUE}
activity_by_date <- with(data, aggregate(steps~interval + datetype, data, mean, na.rm = TRUE))
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```
